---
title: "Get started"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Get started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE,
                      comment  = "#>")
```

The package `rgeoservices` is a R client streamlining access to the services provided by the French National Institute of Geographic and Forest Information (IGN) through its open access [Geoplatform](https://geoservices.ign.fr/documentation/services/services-geoplateforme). This Geoplatform can be used to query the IGN BD TOPO® database to retrieve geographical information.

**Note:** The IGN Geoplatform provides information **only for French locations**.


Let's load and attach the package `rgeoservices`.

```{r setup}
## Attach 'rgeoservices' package ----
library(rgeoservices)
library(mapview)
library(ggplot2)
```


## The Geocoding service

The IGN Geoplatform provides the [Geocoding service](https://geoservices.ign.fr/documentation/services/services-geoplateforme/geocodage). This service can be used in two different ways:

- to retrieve geographical coordinates and administrative information from an address (or a point of interest)
- to retrieve an address (or a point of interest) from geographical coordinates.

### Retrieve geographical coordinates

User can retrieve geographical coordinates and administrative information (postal code, INSEE code, city, etc.) from any French address (postal address, street, city) or any point of interest (POI) such as city hall, museum, public garden, lake, etc. This service is implemented in the [`gs_get_coordinates()`](https://ahasverus.github.io/rgeoservices/reference/gs_get_coordinates.html) function of `rgeoservices`.


#### Postal address

Let's try to retrieve geographical coordinates of a **postal address**.

```{r 'get_coordinates_address', eval=FALSE}
## Retrieve coordinates from an address ----
x <- gs_get_coordinates(
  query = "5 rue de l'Ecole de Médecine", 
  index = "address",
  limit = 10
)
```

```{r 'print_coordinates_address', echo=FALSE}
load("chunks/get_coordinates_address.rda")

get_coordinates_address$"score" <- round(as.numeric(get_coordinates_address$"score"), 5)
knitr::kable(get_coordinates_address)
```

The argument `index` is used to indicate to search from an address (`index = 'address'`) or a point of interest (`index = 'poi'`). Depending on the value of this argument, some columns can be `NA`.

User can filter results by using additional parameters:

- `postcode` or `city` to specify the **postal code** or the **city name** associated to the address
- `type` to specify the precision of the address, e.g. `housenumber`, `street`, `locality`, or `municipality`


```{r 'get_coordinates_address_filter', eval=FALSE}
## Retrieve coordinates from an address (with filters) ----
x <- gs_get_coordinates(
  query    = "5 rue de l'Ecole de Médecine", 
  index    = "address",
  limit    = 10,
  postcode = "34000",
  type     = "housenumber"
)
```

```{r 'print_coordinates_address_filter', echo=FALSE}
load("chunks/get_coordinates_address_filter.rda")

get_coordinates_address_filter$"score" <- round(as.numeric(get_coordinates_address_filter$"score"), 5)
knitr::kable(get_coordinates_address_filter)

x <- get_coordinates_address_filter
```

Using these filters can help getting better results.

Let's display the result on an interactive map.

```{r 'map_coordinates_address_filter'}
## Convert to spatial object ----
x <- sf::st_as_sf(
  x      = x,
  coords = c("longitude", 'latitude'),
  crs    = sf::st_crs(4326)
)

## Setup 'mapview' ----
mapviewOptions(
  basemaps           = "OpenStreetMap",
  legend             = FALSE,
  layers.control.pos = "topright"
)

## Interactive map of the result ----
mapview(
  x           = x, 
  color       = "white", 
  col.regions = "red", 
  cex         = 10
)
```


#### City

Let's try to retrieve geographical coordinates of a **city**.

```{r 'get_coordinates_city', eval=FALSE}
## Retrieve coordinates from a city ----
x <- gs_get_coordinates(
  query    = "Montpellier", 
  index    = "address",
  limit    = 10,
  postcode = "34000",
  type     = "municipality" 
)
```

```{r 'print_coordinates_city', echo=FALSE}
load("chunks/get_coordinates_city.rda")

get_coordinates_city$"score" <- round(as.numeric(get_coordinates_city$"score"), 5)
knitr::kable(get_coordinates_city)

x <- get_coordinates_city
```

Let's map the result.


```{r 'coordinates_city_to_sf'}
## Convert to spatial object ----
x <- sf::st_as_sf(
  x      = x,
  coords = c("longitude", 'latitude'),
  crs    = sf::st_crs(4326)
)
```

```{r 'get_gadm', eval=FALSE}
## Download static map of France (GADM) ----
gadm_fra0 <- geodata::gadm(
  country = "France",
  level   = 0,
  path    = "."
)
```

```{r 'get_gadm_fake', echo=FALSE}
gadm_fra0 <- rgeoservices:::gadm_fra0
```

```{r 'map_coordinates_city'}
ggplot() + 
  geom_sf(data = gadm_fra0) + 
  geom_sf(data = x, size = 4, col = "red") + 
  theme_bw()
```


### Retrieve address or POI

...


## The Altimetry service

...

## The Route planner service

...

## The Isochrone service

...

## The Isodistance service

...


